---
title: "Pigs in a bucket combined analyses"
author: "David S, Mason"
date: "6/19/2021"
output:
  pdf_document: default
  html_document: default
---

---
### Garbage can analyses

This was an experiment designed to explore how carrion biomass and vertebrate scavenegers modulate decomposition rates. Individual pig carcasses were exposed to other decomposing carrion at increasing levels of biomass with and without vertebrate exclusion. After a time period, these individual carcasses werer returned to  This is a factorial design crossing exposure biomass with vertebrate exclusion. There are six observations, one for each combination of carrion biomass and vertebrate exclusion (i.e., no replication). 

The permutation based anova essentially rerranges the values of variables many times to generate a distribution of the test statistic. This test statistic is F for anova (or pseudo-f here), which is a ratio of between and among group variation. We need an approach that fits our data and goals: a nonparametric, permutation-based ancova-like analysis fit for saturated models.Section 2.6 of the supporting document for the lmPerm package provides such an anlysis (Wheeler 2016). The result of this analysis is significant for both treatments and their interaction. 

outstanding question: 

How do I choose between lmp vs aovp? The results do not agree.

## Load packages and bring in the data
```{r setup, echo=FALSE}
d.tmp <- read.csv("Data/garbage_can.csv")

library(tidyverse)
library(RVAideMemoire)
library(simpleboot)
library(sm)
library(lmPerm)
library(multcomp)
library(car)
```
## Calculate the change in biomass
```{r}
d <- d.tmp %>% 
	mutate(LOST_BIO = abs(PIG_FINAL-PIG_START))
```
## Final model using ancova with permutation
```{r}
mod3 <- lmp(LOST_BIO~FENCE*EXPOSURE, data=d)
summary(mod3) 
summary(mod3)$r.squared
Anova(mod3)

# Fence increased biomass lost by 1.19 kg
# 1 kg increase in biomass exposure increased biomass lost by 0.003 kg
# That slope is 0.001 higher in Fence (0.004 kg) 

sem <- function(x, na.rm=FALSE){
		out <- sd(x, na.rm = na.rm)/sqrt(length(x))
		return(out)
}

d %>% group_by(FENCE) %>% 
	summarise(mean = mean(LOST_BIO),
						sd = mean(LOST_BIO),
						se = sem(LOST_BIO))

d %>% group_by(EXPOSURE) %>% 
	summarise(mean = mean(LOST_BIO),
						sd = mean(LOST_BIO),
						se = sem(LOST_BIO))
```

## Visualize the model
```{r, echo=FALSE}

levels(d$FENCE) <- c("Fenced","Open")

ggplot(d,aes(y=LOST_BIO,x=log(EXPOSURE),color=FENCE,fill=FENCE))+
	geom_point(size = 6.5,shape=21,stroke=2)+
	scale_fill_manual(values=c("#00AFBB", "#E7B800"))+
	scale_color_manual(values=c("black","black"))+
	stat_smooth(method="lm",se=FALSE,show.legend = FALSE)+
	scale_x_continuous(name="Biomass treatment (kg)",
										 breaks=c(4.007333,5.991465,7.377759),
										 labels=c("55","400","1600"))+
	ylab("Lost biomass (kg)")+
	theme_classic()+
	theme(legend.title = element_blank())+
	theme(legend.position = c(0.17, 0.92))+
	theme(text = element_text(size = 25))+
	theme(axis.title.x = element_text(face="bold"))+
	theme(axis.title.y = element_text(face="bold", vjust=0.7))+
		 guides(fill=guide_legend(
                 keyheight=0.4,
                 default.unit="inch"))

ggsave('garbage_can.jpg', width = 6, height = 6, dpi = 300)
```

## The y axis is the absolute value in change in biomass, and the x axis is log-transformed. The labels are back-transformed for interpretability. 

### Fly trapping analyses

In the original field experiment (i.e., not the additional carcassses referenced
in the analysis above), the researchers trapped flies to demonstrate the difference
in abundance days/weeks sost-deployment associated with biomass and the exclusion
of vertebrate scavengers (Fenced). 

I devloped a model with biomass (continuous) and exclusion (categorical) as fixed
effects. In this experiment, we are not interested in the effect of time on fly
abundance, beyond the need to account for variance in the model. Thus, I set 
date as a random effect. 

## Load packages and bring in the data
```{r}
library(tidyverse)
library(lme4)
library(lubridate)
library(car)
library(DHARMa)
library(emmeans)
library(MuMIn)

d <- read.csv("Data/flies.csv")
head(d)
summary(d)

d$Date <- paste(d$Year, d$Month, d$Day, sep="-") %>% ymd() %>% as.Date()
```
## Visualize the data
```{r}
ggplot(d, aes(y=AVGrate, x=Date, color=Fence))+
	geom_point()+
	facet_wrap(~Biomass)
```
## Model the interacton between fencing and the relationship between biomass 
and flies trapped
```{r}
d.tmp <- filter(d, !Date=='2016-07-13')

# Simple model
mod3 <- lm(AVGrate ~ Date * Biomass * Fence, data = d.tmp)
plot(mod3)
hist(mod3$residuals, breaks=10)
anova(mod3)

# Mixed effects model
mod4 <- lmer(AVGrate ~ Biomass * Fence + (1|Date), data = d.tmp)

mod4.sim <- simulateResiduals(fittedModel = mod4, n = 250)
plot(mod4.sim)

Anova(mod4)
```
## Model results
```{r}
## emmeans, correcting for bias (only needed for GLMMs) by calculating the total SD of the random effect variance components
modSD <- VarCorr(mod4) %>% as.data.frame() %>% summarize(totSD=sum(vcov)) %>% mutate(totSD=sqrt(totSD))

emmeans(mod4, pairwise~Fence, type="response", bias.adj=T, sigma=modSD$totSD)
# flies caught per minute is 0.204 (SE = 0.0805) higher in fence than control
# fence mean is 0.596 (LCL = 0.231, UCL = 0.961)
# open mean is 0.392 (LCL = 0.027, UCL = 0.758)
emtrends(mod4, pairwise~Fence, var = 'Biomass') 
# each 1 kg of biomass increased mean flys caught per minute by 0.000451
# (SE = 0.000139) for fence in comparison to control.
# fence per unit (kg) increase in flies is 0.000804 (LCL = 0.000603, UCL = 0.001004)
# open per unit (kg) increase in flies is 0.000353 (LCL = 0.000152, UCL = 0.000554)

r.squaredGLMM(mod4) # fixed efects = 0.557324, entire model = 0.759619
VarCorr(mod4) # random effects = 0.22769, unexplained variance = 0.24820
```
## Visualize the model
```{r}
d.NA <- d.tmp[!is.na(d.tmp$AVGrate),]

sem <- function(x, na.rm = FALSE) {
  out <-sd(x, na.rm = na.rm)/sqrt(length(x))
  return(out)
}

d.means <- d.NA %>% 
							group_by(Biomass,Fence) %>% 
							summarise(mean = mean(AVGrate*60),
								sd = sd(AVGrate*60),
								se = sem(AVGrate*60))

# Get values for back-transformed axis labels
log(55) # 4.007333
log(130) # 4.867534
log(400) # 5.991465
log(800) # 6.684612 
log(1600) # 7.377759 

levels(d.means$Fence) <- c("Fenced","Open")

ggplot(d.means, aes(x=log(Biomass),y=mean,color=Fence,fill=Fence))+
	geom_errorbar(aes(ymin=mean-se,ymax=mean+se),
								position = position_dodge(width = 0.5),color='black', width=0.2)+
	geom_point(size=6.5,shape=21,stroke=2,position=position_dodge(width = 0.5),color='black')+
	scale_color_manual(values=c("#00AFBB", "#E7B800"))+
	scale_fill_manual(values=c("#00AFBB", "#E7B800"))+
	scale_x_continuous(name="Biomass treatment (kg)",
										 breaks=c(4.007333,4.867534,5.991465, 6.684612, 7.377759),
										 labels=c("55", "130", "400", "800", "1600"))+
	scale_y_continuous(name="Mean flies per hour",
										 breaks=c(0.0*60,0.25*60,0.50*60,0.75*60,1.00*60,1.25*60,
										 				 1.5*60,1.75*60,2*60),
										 limits = c(0,2*60))+
	theme_classic()+
	theme(legend.title = element_blank())+
	theme(legend.position = c(0.17, 0.92))+
	theme(text = element_text(size = 25))+
	theme(axis.title.x = element_text(face="bold"))+
	theme(axis.title.y = element_text(face="bold", vjust=0.7))+
	 guides(fill=guide_legend(
                 keyheight=0.4,
                 default.unit="inch"))

ggsave('flies_drop_first.jpg', width = 6, height = 6, dpi = 300)
```

